CS6213 Project Report

### 2021-05-07 report/paper.tex
WIP summary

Due to time constraints, we had to significantly reduce our ambitions.
%
In the end, we evaluated only the most widely used implementation, \etcd, and not nearly to the extent we had initially hoped.
%
Nonetheless, our experience has been instructive and leads us to believe there is potential in continuing this line of work.

>>>>

Due to time constraints, we had to reduce our ambitions.
%
In the end, we evaluated only the most widely used implementation, \etcd, and not to the extent we had initially hoped.
%
Nonetheless, our experience has been instructive and leads us to believe there is potential in continuing this line of work.

\section{Project Summary}

\begin{itemize}
    \item We combined three existing TLA specifications (due to Diego Ongaro, Daniel Ricketts, and Brandon Amos and Hangcheng Zhang) into a reference specification of Raft with one-at-a-time membership changes.

    \item We started developing a prototype testing framework on top of \etcd to simulate execution traces generated by TLC.

    \item We noticed that model checking performance is low.
    %
    With the default TLC configuration, on a model with 3 nodes, it took 2 minutes before the first entry was added to a log, when running TLC on our laptops.
    %
    Entries began to be committed only much later.

    \item Hoping for a better model checking experience, we adapted our specification for use with the Apalache symbolic model checker.
    %
    This required adding type annotations to the TLA specification and slightly rewriting some operators to fit into the subset of TLA that can be automatically translated to SMT.
    %
    Apalache is alpha software, so we ran into a large number of bugs.
    %
    Sadly, for our specification, Apalache symbolic model checking was orders of magnitude slower than TLC.

    \item Disappointed with Apalache's performance, we reverted to using TLC to generate Raft execution traces.

    \item In the process, we discovered (low impact) errors in the
    specification of membership changes due to Amos and Zhang and in the
    properties stated by Daniel Rickkets.
    %
    This suggests that their specifications have been model-checked only to a shallow extent.
    %
    Given the effort we spent to obtain reasonable performance out of TLC, this is understandable.
